\documentclass[11pt]{article}
\usepackage[letterpaper,margin=0.9in]{geometry}
\usepackage[hidelinks]{hyperref}
\usepackage{enumitem}
\usepackage{titlesec}
\usepackage{tabularx}

\setlist[itemize]{leftmargin=1.2em,itemsep=3pt,topsep=3pt}
\titleformat{\section}{\large\bfseries}{}{0em}{}[\titlerule]

\newcommand{\entry}[4]{
  \noindent\begin{tabularx}{\textwidth}{@{}Xr@{}}
    \textbf{#1} & #2 \\
    \textit{#3} & \textit{#4}
  \end{tabularx}\vspace{2pt}
}

\begin{document}

\begin{center}
  {\LARGE \textbf{Piter Z. Garcia Bautista}}\\[4pt]
  Rochester, NY \textbar\ +1 (646) 288-3300 \textbar\ \href{mailto:pzg8794@rit.edu}{pzg8794@rit.edu}\\
  \href{https://linkedin.com/in/garciapiterz}{linkedin.com/in/garciapiterz} \textbar\
  \href{https://github.com/pzg8794}{github.com/pzg8794} \textbar\
  \href{https://garciapiterz.com}{garciapiterz.com}
\end{center}

\section{Research Profile}
Graduate researcher with a strong foundation in machine learning methods, statistical learning theory, and data-driven decision support. Background includes fairness-aware ML for high-stakes settings, reproducible evaluation frameworks, bandit and sequential decision-making algorithms, and applied ML across healthcare, bioinformatics, and network routing domains. Current Graduate Assistant work (Quantum and AI Research, RIT) emphasizes principled model evaluation, statistical rigor, and method reliability across heterogeneous and partially observed data environments. Interested in pushing ML methodology forward --- building models that generalize robustly and whose failure modes are understood and bounded.

\section{Research Interests}
Machine learning theory and methods; statistical learning and model evaluation; fairness-aware and robust ML; bandit and sequential decision-making; reproducible benchmarking; applied ML in healthcare, bioinformatics, and network systems.

\section{Education}

\entry{Rochester Institute of Technology}{Expected Aug 2026}{Master of Science, Data Science --- GPA 3.9+}{Rochester, NY}
\begin{itemize}
  \item Concentration: Bioinformatics \& Artificial Intelligence.
  \item Relevant coursework: Data-Driven Knowledge Discovery (ISTE-780), Neural Networks \& Deep Learning, Applied Data Science (DSCI-601), Bioinformatics Algorithms (BIOL-630), Software Engineering for Data Science (DSCI-644), Foundations of Data Science (DSCI-633).
\end{itemize}

\entry{University of Rochester, Warner School of Education}{In progress}{Graduate study, Teaching Computer Science K--12 (M.S. track)}{Rochester, NY}
\begin{itemize}
  \item NSF Robert Noyce Scholar (\$30,000 full funding); Advanced Certificate in Leadership in Disability and Inclusive Practices.
\end{itemize}

\entry{Rochester Institute of Technology}{2015}{Master of Science, Computer Science}{Rochester, NY}
\begin{itemize}
  \item Concentrations: Artificial Intelligence, Big Data Analytics, Computer Graphics.
  \item Graduate research: \textit{Big Data Medical Diagnosis: A Multi-Method Approach} --- ML ensemble methods for clinical classification.
\end{itemize}

\entry{Farmingdale State College}{2012}{Bachelor of Science, Computer Engineering}{Farmingdale, NY}
\begin{itemize}
  \item Dual degree: Electrical Engineering \& Computer Engineering Technology.
  \item Honors: CSTEP Award \& Merit; Dual BS Scholarship.
\end{itemize}

\section{Research and Professional Experience}

\entry{Rochester Institute of Technology, Software Engineering Department}{Fall 2025 -- Present}{Graduate Assistant, Quantum and AI Research}{Rochester, NY}
\begin{itemize}
  \item Build reproducible ML evaluation pipelines across local, Colab, and GCP environments with structured logging, hyperparameter tracking, and statistical validation.
  \item Develop bandit and sequential decision-making systems (UCB, neural-UCB, contextual bandits) with rigorous evaluation frameworks comparing algorithmic variants under controlled conditions.
  \item Produce manuscript-ready technical analyses on ML method performance, including evaluation under distribution shift, partial observability, and adversarial settings.
  \item Design fairness-aware evaluation protocols that measure both predictive performance and equity metrics across heterogeneous subpopulations.
\end{itemize}

\entry{Rochester Institute of Technology}{2024 -- 2025}{Equitable ML Research --- ISTE-780 Project (Data-Driven Knowledge Discovery)}{Rochester, NY}
\begin{itemize}
  \item Pioneered first systematic application of fairness assessment to RNA secondary structure prediction algorithms (EQUITAS framework), revealing critical equity violations (11.1\% equity score) across demographic proxies.
  \item Developed modular ML evaluation codebase (900+ lines) implementing reproducible fairness-aware algorithm comparison with automated equity-metric reporting.
  \item Achieved statistical significance ($p < 0.01$) in bias detection across multiple algorithmic variants using SHAP and disparity-impact analysis.
\end{itemize}

\entry{Rochester Institute of Technology}{2025 -- 2026}{Computational Biology Research (BIO614 and BIO550)}{Rochester, NY}
\begin{itemize}
  \item Applied ML and statistical methods to RNA structure prediction, including evaluation of probabilistic models against classical dynamic-programming baselines.
  \item Conducted RNA-seq differential-expression analysis with reproducible preprocessing, QC validation, and statistical-significance reporting.
\end{itemize}

\entry{VEDADATA Inc.}{Sep 2022 -- Aug 2024}{Data Solutions Engineer}{New York, NY}
\begin{itemize}
  \item Designed scalable Python and pandas ML workflows for ingestion, feature engineering, and statistical diagnostics in production environments.
  \item Implemented structured quality checks and anomaly-detection pipelines strengthening production data reliability.
\end{itemize}

\entry{VIOME Inc.}{Jan 2021 -- Sep 2022}{AI Data Solutions Engineer}{New York, NY}
\begin{itemize}
  \item Developed healthcare ML workflows for feature engineering, model experimentation, and evaluation --- predictive health applications in AWS production environments.
  \item Built reproducible model-evaluation infrastructure with monitoring components to detect performance degradation under distribution shift.
\end{itemize}

\entry{Samasta Health Foundation}{Mar 2017 -- Jan 2021}{Computer \& Data Science Consultant (pro bono)}{Remote}
\begin{itemize}
  \item Designed AI-driven healthcare solutions emphasizing equity and accessibility for underserved populations.
\end{itemize}

\section{Selected Technical Writing}
\begin{itemize}
  \item \textit{Fairness-Aware Bandits for Network Routing in Quantum and Clinical Settings} (DSCI601 project, manuscript-style) --- bandit learning with equity constraints for high-stakes sequential decisions.
  \item \textit{Equitable Bioinformatics: Enhancing Diagnostic Decision-Making through RNA and Biomarker Data} (ISTE-780 Project Report) --- systematic fairness assessment applied to ML-driven diagnostics.
  \item \textit{BIO614-FinalProjectProposal} (Overleaf) --- formal computational methods manuscript with rigorous benchmarking and reproducible evaluation.
\end{itemize}

\section{Awards \& Recognition}
\begin{itemize}
  \item NSF Robert Noyce Teacher Scholarship (2024--2026) --- full funding for M.S. in Teaching Computer Science K--12.
  \item MS International Scholarship (2012--2015) --- MESCYT \& RIT, awarded for academic excellence.
  \item Honorable Mention Technology Award (2012) --- NYS STEP/CSTEP, 2nd place capstone project.
  \item IEEE Alumni Member (2009--Present) --- Member No. 90613000.
\end{itemize}

\section{Technical Competencies}
\textbf{Programming:} Python, R, Java, C++, C\#, SQL, MATLAB, JavaScript, Bash\\
\textbf{Machine Learning:} PyTorch, TensorFlow, scikit-learn, XGBoost; bandit/RL (UCB, contextual bandits, Thompson sampling); fairness-aware ML; model evaluation and uncertainty quantification; SHAP analysis\\
\textbf{Data and Statistics:} pandas, NumPy, SciPy, statistical modeling, hypothesis testing, reproducible benchmarking\\
\textbf{Infrastructure:} Git/GitHub, Docker, Linux, GCP, AWS, LaTeX

\section{Selected Fit for KTH Machine Learning Position}
\begin{itemize}
  \item Strong theoretical and applied ML background spanning supervised, sequential decision-making, and fairness-aware methods, with emphasis on rigorous evaluation methodology.
  \item Hands-on experience designing and running controlled ML experiments with statistical validation --- directly applicable to research that advances ML methodology.
  \item Demonstrated commitment to reproducibility: all research projects include structured evaluation pipelines, logging, and statistical-significance reporting.
  \item Deep interest in understanding when and why ML methods work or fail, with particular focus on robustness, equity, and reliability in high-stakes applications.
\end{itemize}

\end{document}
